Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae

### 2021-05-02 thesis_chapter_3.Rmd
results AJ comments

The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (Figure 2A). When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen; rather, there is only a slight variation. The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation (Figure 2B). The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well the final location of the kelp particles being further from the release site. When comparing the different hydrodynamic and wind drag simulations to each other, only slight variation in mean trajectories was evident. In all simulations particles flowed in a north-westward direction.

The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (autumn). Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag traveled further northward compared to the passive particles (see Figure 2C-D). The mean trajectories of kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles. Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles traveled less distance compared to kelp particles with both wind and hydrodynamic drag. Differences in mean trajectories between kelp and passive particles were also evident in the July (winter) simulations (see Figure 2E-F). Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore; resulting in entrainment within an eddy. Similar behavior is evident when comparing passive particles with kelp particles exposed to both hydrodynamic and wind drag. In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles traveled initially southward and then westward offshore and also becoming entrained in an eddy. In the October (Spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 7F-G). Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag ended a greater distance from the release site compared to the passive particles. This was also evident when comparing kelp particles with different cross-sectional areas and exposure to wind and hydrodynamic drag with passive particles; with the end location a further from the release location. In both the April and October simulations all particles flowed north-westward, while in winter the mean trajectory resulted in entrainment within an eddy.

@@

Across all simulations particles flowed in a north-westward direction. The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids. In addition the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 3). Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.

In April and across all simulations, particles flowed in a north-westward direction (see Figure 4). The density plots that considered any form of drag showed narrow, higher density along the mean trajectory compared to passive simulation. In addition, simulations that considered any form of drag got entrained within an eddy further north-westward from the end location of passive particles. Only slight variations across cross-sectional areas and drag exposure scenarios are evident. Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in a eddy located westward of the release site compared to passive particles (see Figure 5). Variation in density along the mean trajectory and eddy for all kelp particles is evident. Kelp particles with only hydrodynamic drag across cross-sectional areas showed lower density of particles along the mean trajectory compared to kelp particles with both wind and hydrodynamic drag. October simulations showed no differences in density of particles along the mean trajectory when comparing passive and kelp particles. Only slight variations in density and trajectory can be seen with no entrainment (see Figure 6). Kelp particles with any cross-sectional area and drag exposure scenario flowed further offshore when compared to passive particles.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end run time for scenarios 14 - 26, for April 2018 (autumn). The plots are regular hexagons within the density of particles were calculated. The mean trajectories for each scenario are also plotted for comparison. The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end run time for scenarios 27 - 39, for July 2018 (winter). The plots are regular hexagons within the density of particles were calculated. The mean trajectories for each scenario are also plotted for comparison. The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end run time for scenarios 40 - 52, for October 2018 (spring). The plots are regular hexagons within the density of particles were calculated. The mean trajectories for each scenario are also plotted for comparison. The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}

@@

There were significant differences among the median distances traveled for all particles in each simulation (Figure 4; p < 0.05). The simulations that considered for 100% hydrodynamic drag traveled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios. Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations. When comparing distances between passive and kelp particles for April, significant differences are evident. (Figure 11, simulations 17-29; p < 0.05). Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas traveled shorter distance across compared to other kelp particles. The shortest distance traveled was the scenario with the maximum cross-sectional area and drag exposure scenarios. No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; p > 0.05). In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; p < 0.05). Kelp particles with only 90% hydrodynamic drag across cross-sectional areas traveled shorter distance compared to all other particles. As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure traveled the shortest distance compared to all other particles.

@@

```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 7: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for scenarios 1 - 13, for January 2018 (summer). The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers. The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=7, fig.width=5}

>>>>

The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (Figure 2A). When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen; rather, there is only a slight variation. The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation (Figure 2B). The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well the final location of the kelp particles being further from the release site. When comparing the different hydrodynamic and wind drag simulations to each other, only slight variation in mean trajectories was evident. In all simulations, particles flowed in a north-westward direction.

The mean trajectories of kelp particles were different compared to the mean trajectory of passive particles for April (autumn). Mean trajectories for kelp particles with different cross-sectional areas exposed to only hydrodynamic drag traveled further northward compared to the passive particles (see Figure 2C-D). The mean trajectories of kelp particles with different cross-sectional area types and a combination of wind and hydrodynamic drag flowed further north-westward compared to the passive particles. Across all simulations for April, the end locations were different, and hydrodynamic only kelp particles traveled less distance compared to kelp particles with both wind and hydrodynamic drag. Differences in mean trajectories between kelp and passive particles were also evident in the July (winter) simulations (see Figure 2E-F). Mean trajectories of passive particles flowed north-westward, while kelp particles across cross-sectional areas and exposed to only hydrodynamic drag initially flowed in a southerly direction and then flowed westward offshore, resulting in entrainment within an eddy. Similar behavior is evident when comparing passive particles with kelp particles exposed to both hydrodynamic and wind drag. In the combined hydrodynamic and water drag scenarios, and across cross-sectional areas; kelp particles traveled initially southward and then westward offshore and also becoming entrained in an eddy. In the October (Spring) simulations, differences in end locations were evident between passive and kelp particles across cross-sectional areas and drag scenarios (see Figure 7F-G). Mean Trajectory of kelp particles with different cross-sectional areas and exposed to only hydrodynamic drag ended a greater distance from the release site compared to the passive particles. This was also evident when comparing kelp particles with different cross-sectional areas and exposure to wind and hydrodynamic drag with passive particles, with the end location a further from the release location. In both the April and October simulations all particles flowed north-westward, while in winter the mean trajectory resulted in entrainment within an eddy.

@@

Across all simulations, particles flowed in a north-westward direction. The density plots for simulations that considered any form of drag showed a higher density of particles along the mean trajectory path compared to the passive simulation which has an almost even density of particles across grids. In addition, the density plots show that the particles in simulations that considered any form of drag got entrained in a vorticity while particles in the passive simulation deflected away (Figure 3). Comparison of cross-sectional areas and different drag exposure scenarios show no differences in mean trajectory and spatial patterns in density, with only slight variations across types.

In April and across all simulations, particles flowed in a north-westward direction (see Figure 4). The density plots that considered any form of drag showed narrow, higher density along the mean trajectory compared to passive simulation. In addition, simulations that considered any form of drag got entrained within an eddy further north-westward from the end location of passive particles. Only slight variations across cross-sectional areas and drag exposure scenarios are evident. Simulations for July showed that the flow was also north-westward, but all kelp particles across cross-sectional areas and drag scenarios show high density of particles entrained in an eddy located westward of the release site compared to passive particles (see Figure 5). Variation in density along the mean trajectory and eddy for all kelp particles is evident. Kelp particles with only hydrodynamic drag across cross-sectional areas showed lower density of particles along the mean trajectory compared to kelp particles with both wind and hydrodynamic drag. October simulations showed no differences in density of particles along the mean trajectory when comparing passive and kelp particles. Only slight variations in density and trajectory can be seen with no entrainment (see Figure 6). Kelp particles with any cross-sectional area and drag exposure scenario flowed further offshore when compared to passive particles.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 4: Comparison of density plots at the end of the run for scenarios 14 - 26, for April 2018 (autumn). The plots are regular hexagons within the density of particles were calculated. The mean trajectories for each scenario are also plotted for comparison. The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 5: Comparison of density plots at the end of the run for scenarios 27 - 39, for July 2018 (winter). The plots are regular hexagons within the density of particles were calculated. The mean trajectories for each scenario are also plotted for comparison. The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 6: Comparison of density plots at the end of the run for scenarios 40 - 52, for October 2018 (spring). The plots are regular hexagons within the density of particles were calculated. The mean trajectories for each scenario are also plotted for comparison. The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=8, fig.width=8}

@@

There were significant differences among the median distances traveled for all particles in each simulation (Figure 4; *p* < 0.05). The simulations that considered for 100% hydrodynamic drag traveled a shorter distance than particles for simulations which considered a combination of hydrodynamic and wind drag scenarios. Comparison between simulations that considered hydrodynamic and wind drag show no significant differences with only slight variations among simulations. When comparing distances between passive and kelp particles for April, significant differences are evident. (Figure 11, simulations 17-29; *p* < 0.05). Kelp particles with only 90% hydrodynamic and 10% wind drag across cross-sectional areas traveled shorter distance across compared to other kelp particles. The shortest distance traveled was the scenario with the maximum cross-sectional area and drag exposure scenarios. No significant differences are evident for the July simulations, with only slight variation across simulations (see Figure 11, simulations 30-42; *p* > 0.05). In October, the plot shows significant differences when comparing both passive and kelp particles (see Figure 11, simulations 43-55; *p* < 0.05). Kelp particles with only 90% hydrodynamic drag across cross-sectional areas traveled shorter distance compared to all other particles. As seen in the April simulations, the kelp particles with the maximum cross-sectional area and drag exposure traveled the shortest distance compared to all other particles.

@@

```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 7: Boxplots of total distance travelled by all particles from the release site over the course of the simulations for scenarios 1 - 52. The *x*-axis represents the scenario number, and the *y*-axis is the distance travelled in kilometers. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers. The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=7, fig.width=5}
